---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library("purrr") # prevent `purrr` load message by `furrr`
devtools::load_all()
```

# tidier

<!-- badges: start -->

[![CRAN status](https://www.r-pkg.org/badges/version/tidier)](https://CRAN.R-project.org/package=tidier)
<!-- badges: end -->

`tidier` package provides '[Apache Spark](https://spark.apache.org/)' style window aggregation for R dataframes via '[mutate](https://dplyr.tidyverse.org/reference/mutate.html)' in '[dplyr](https://dplyr.tidyverse.org/index.html)' flavour.

## Example

**Create a new column with average temp over last seven days in the same month**.

```{r}
set.seed(101)
airquality |>
  # create date column
  dplyr::mutate(date_col = lubridate::make_date(1973, Month, Day)) |>
  # create gaps by removing some days
  dplyr::slice_sample(prop = 0.8) |> 
  # compute mean temperature over last seven days in the same month
  tidier::mutate(avg_temp_over_last_week = mean(Temp, na.rm = TRUE),
                 .order_by = Day,
                 .by       = Month,
                 .frame    = c(lubridate::days(7), # 7 days before current row
                               lubridate::days(-1) # do not include current row
                               ),
                 .index    = date_col
                 )
```

## Features

- `mutate` supports
  - `.by` (group by),
  - `.order_by` (order by),
  - `.frame` (endpoints of window frame),
  - `.index` (identify index column like date column),
  - `.complete` (whether to compute over incomplete window).
  
- `mutate` automatically uses a future backend (via [`furrr`](https://furrr.futureverse.org/)).

## Motivation

This implementation is inspired by Apache Spark's [`windowSpec`](https://spark.apache.org/docs/3.2.1/api/python/reference/api/pyspark.sql.Column.over.html?highlight=windowspec) class with [`rangeBetween`](https://spark.apache.org/docs/3.2.1/api/python/reference/api/pyspark.sql.WindowSpec.rangeBetween.html) and [`rowsBetween`](https://spark.apache.org/docs/3.2.1/api/python/reference/api/pyspark.sql.WindowSpec.rowsBetween.html).

## Ecosystem

1.  [`dbplyr`](https://dbplyr.tidyverse.org/) implements this via [`dbplyr::win_over`](https://dbplyr.tidyverse.org/reference/win_over.html?q=win_over#null) enabling [`sparklyr`](https://spark.rstudio.com/) users to write window computations. Also see, [`dbplyr::window_order`/`dbplyr::window_frame`](https://dbplyr.tidyverse.org/reference/window_order.html?q=window_fr#ref-usage).

2.  [`tidypyspark`](https://talegari.github.io/tidypyspark/_build/html/index.html) python package implements `mutate` style window computation API for pyspark.

## Installation

-   dev: `remotes::install_github("talegari/tidier")`
-   cran: `install.packages("tidier")`

## Acknowledgements

`tidier` package is deeply indebted to two amazing packages and people behind it.

1.  [`dplyr`](https://cran.r-project.org/package=dplyr):

```{=html}
<!-- -->
```
    Wickham H, François R, Henry L, Müller K, Vaughan D (2023). _dplyr: A
    Grammar of Data Manipulation_. R package version 1.1.0,
    <https://CRAN.R-project.org/package=dplyr>.

2.  [`slider`](https://cran.r-project.org/package=slider):

```{=html}
<!-- -->
```
    Vaughan D (2021). _slider: Sliding Window Functions_. R package
    version 0.2.2, <https://CRAN.R-project.org/package=slider>.
