---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

devtools::load_all()
```

# tidier

<!-- badges: start -->

[![CRAN status](https://www.r-pkg.org/badges/version/tidier)](https://CRAN.R-project.org/package=tidier) [![R-CMD-check](https://github.com/talegari/tidier/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/talegari/tidier/actions/workflows/R-CMD-check.yaml)

<!-- badges: end -->

`tidier` package provides [Apache Spark](https://spark.apache.org/) style window aggregation for R dataframes via [mutate](https://dplyr.tidyverse.org/reference/mutate.html) in [dplyr](https://dplyr.tidyverse.org/index.html) flavor.

## Example

**Create a new column with average temp over last seven days in the same month**.

```{r}
set.seed(101)
airquality |>
  # create date column
  dplyr::mutate(date_col = lubridate::make_date(1973, Month, Day)) |>
  # create gaps by removing some days
  dplyr::slice_sample(prop = 0.8) |> 
  # compute mean temperature over last seven days in the same month
  tidier::mutate(avg_temp_over_last_week = mean(Temp, na.rm = TRUE),
                 .order_by = date_col,
                 .by       = Month,
                 .frame    = range_between(lubridate::days(7), # 7 days before current row
                                           lubridate::days(-1) # do not include current row
                                           )
                 )
```

## Features

- `mutate` supports
  - `.by` (group by),
  - `.order_by` (order by),
  - `.frame` (window frame defined by `rows_between` or `range_between`),
  - `.complete` (whether to compute over incomplete window).
- `tidier::mutate` is single-threaded. For heavy parallelization across many groups, users can combine `tidyr::nest()` with parallel map (e.g. `furrr::future_map()`) and `tidyr::unnest()`.

## Motivation

This implementation is inspired by Apache Spark's `windowspec`, `rows between` and `range between`.

## Installation

- dev: `remotes::install_github("talegari/tidier")`
- cran: `install.packages("tidier")`

## Acknowledgements

`tidier` package is deeply indebted to the amazing packages and people behind it.

1.  [`dplyr`](https://cran.r-project.org/package=dplyr):

```         
Wickham H, François R, Henry L, Müller K, Vaughan D (2023). _dplyr: A
Grammar of Data Manipulation_. R package version 1.1.0,
<https://CRAN.R-project.org/package=dplyr>.
```

2.  [`slider`](https://cran.r-project.org/package=slider):

```         
Vaughan D (2021). _slider: Sliding Window Functions_. R package
version 0.2.2, <https://CRAN.R-project.org/package=slider>.
```
